---
title: "NFLBDB Calculating Distance"
output: html_notebook
---

## Overview
This notebook will calculate the distance between O-linemen and the passer after
one of the following events have occurred: a pass has been thrown, a sack
occurred, or a strip sack occurred. If a sack or hit on the quarterback occurred,
and the O-lineman responsible for blocking the player that gets the sack or hit, 
then the distance denied will be 0. If the O-lineman is beaten, then the distance
denied will be equal to the distance between the pass rusher that beat the
O-lineman and the passer. Otherwise, the distance denied will be equal to the
O-lineman and the passer at the time of the pass being thrown or the sack
occurring.

## Import Data

```{r, warning=FALSE, include=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
```

```{r}
setwd("C:\\Users\\samue\\OneDrive\\Desktop\\NFL_BDB")
df <- read.csv("week1_all_data_completely_cleaned.csv")
head(df)
```
## Additional Cleaning of the Table
I need to get rid of some NA values in the table because R will throw errors if these are read.
I will change these values to 0 so that the logic in calculating distance denied is not corrupted.

```{r}
df[is.na(df)] <- 0
```

## Calculate Distance Denied by O-linemen

```{r, warning=FALSE}
# First, create a new column in the dataframe called "Distance Denied"
df$Distance_denied <- NA

# Then, loop through each row in the dataframe
for (i in 1:nrow(df)) {
  # Check if the passer was sacked or hit
  if (df$pff_sackAllowed[i] == 1 || df$pff_hitAllowed[i] == 1) {
    # If the passer was sacked or hit, set the distance to 0
    df$Distance_denied[i] <- 0
  } else {
    # If the passer was not sacked or hit, proceed with calculating the distance
    
    # Extract the x and y coordinates of the passer and the O-lineman
    passer_x <- df$x[df$pff_role == 'Pass' & df$playIdUnique == df$playIdUnique[i]]
    passer_y <- df$y[df$pff_role == 'Pass' & df$playIdUnique == df$playIdUnique[i]]
    olineman_x <- df$x[i]
    olineman_y <- df$y[i]
    
    # Check if the O-lineman was beaten by a pass rusher
    if (df$pff_beatenByDefender[i] == 1) {
      # If the O-lineman was beaten, find the x and y coordinates of the pass rusher who beat him
      pass_rusher_x <- df$x[df$pff_nflIdBlockedPlayer[i] == df$nflId & df$playIdUnique == df$playIdUnique[i]]
      pass_rusher_y <- df$y[df$pff_nflIdBlockedPlayer[i] == df$nflId & df$playIdUnique == df$playIdUnique[i]]
      
      # check to see if the blocked player has data for this play. If the data is missing,
      # then just use o-lineman_x and y data for the positioning to avoid error.
      if (length(pass_rusher_x) == 0) {
        pass_rusher_x <- olineman_x
      }
      if (length(pass_rusher_y) == 0) {
        pass_rusher_y <- olineman_y
      }
      
      # Use the distance formula to calculate the distance between the pass rusher and the passer
      distance <- sqrt((passer_x - pass_rusher_x)^2 + (passer_y - pass_rusher_y)^2)
    } else {
      # If the O-lineman was not beaten, use the distance formula to calculate the distance between the passer and the O-lineman
      distance <- sqrt((passer_x - olineman_x)^2 + (passer_y - olineman_y)^2)
    }
    
    # Save the calculated distance in the "Distance" column of the dataframe
    df$Distance_denied[i] <- distance
  }
}
```

## Export the data
write the dataframe to a csv file

```{r}
write.csv(df, "C:\\Users\\samue\\OneDrive\\Desktop\\NFL_BDB\\distance_denied_week1.csv", row.names=FALSE)
```