---
title: "Distance Denied By O-Lineman Per Play"
output: html_notebook
---

## Overview
This notebook will take the calculations for distance denied by offensive lineman for individual plays, aggregate that data, and then calculate an average distance per play. This final result will allow the offensive linemen to be ranked with each other.

## Import Data

```{r, warning=FALSE, include=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
```

```{r}
setwd("C:\\Users\\samue\\OneDrive\\Desktop\\NFL_BDB")
df <- read.csv("distance_denied_by_Olineman.csv")
head(df)
```

## Aggregate the data and find average distance denied
```{r}
data_grouped <- df %>%
  group_by(displayName)
  
data_summary <- data_grouped %>%
  summarise(
    total_distance = sum(Distance_denied),
    total_plays = n()
  )

# Use the mutate() function to calculate the average distance run per play for each player
data_summary <- data_summary %>%
  mutate(
    average_distance_denied_per_play = total_distance / total_plays
  )

# Print the summary data
print(data_summary)
```

## Export the data
write the dataframe to a csv file

```{r}
write.csv(data_summary, "C:\\Users\\samue\\OneDrive\\Desktop\\NFL_BDB\\distance_denied_per_play_week1.csv", row.names=FALSE)
```
